---
title: "Take-home exercise 1"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This take-home exercise is based on CSV data set associated with a recent paper in the journal *PLOS Biology* on [root-associated microbiomes in rice](https://journals.plos.org/plosbiology/article?id=10.1371/journal.pbio.2003862).
You can download this CSV file to your home directory on the server at `/data/arbuckle1.csv`.

The [original data set](https://datadryad.org/resource/doi:10.5061/dryad.7q7k1/1) had 11,350,506 rows.
I've aggressively filtered this down to a manageable size.
So much so that this data set represents only 3 plants!

1. Load the contents of this CSV file into R and display the numbers of rows and columns.
```{r}

```

2. The column `variable` contains numeric ID's for different soil microbe OTUs ([operational taxonomic units](https://en.wikipedia.org/wiki/Operational_taxonomic_unit)).
I've filtered the original data so that only 66 of these OTUs are present for every root specimen.
Since every row in this table corresponds to an OTU per root specimen, other variables that are defined at the level of specimens are each repeated 66 times!

    Create a new data frame by subsetting this data frame on rows where `variable` equals `4386` (*Leptospiraceae*, a family of spirochete bacteria).

```{r}

```


3. Using your reduced data frame from question (2), create a histogram of `Depth`:
```{r}

```


4. Let's go back to the data frame from question (1).  Some of these columns contain all missing values.  They're sort of annoying, so use `summary` to figure out which columns these are (the number of `NA` values should equal the number of rows), and then use a numeric vector of negative integer indices to exclude these columns in a new data frame.  (Remember that R is 1-indexed, while Python is 0-indexed!)  **Display the first 6 rows of your new data frame.**
```{r}

```

5. Create another data frame by importing the contents of another CSV file, which you can obtain from the server at `/data/soilOTUs.csv`.  Merge the `Class` column from this data frame into your data frame from question (4) using the approach I described in class (you can also find this in the HTML lecture materials).  In other words, you need to apply the `match` function to the vectors `variable` and `OTU`.
```{r}

```
